---
title: "General Social Survey"
author: "Giselle Gonzalez"
date: "Year 2024"
output: 
  html_document:
    number_sections: true
    toc: true
---


```{r include=FALSE,echo=FALSE}
require(tigerstats)
require(tidyverse)
require(foreign)
source("GrabData.R")
```




#  Question 1

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

The data in the dataframe GSSdata is from the 2018 General Social Survey. The first blocks of R-code has selected down a subset of the data to just 16 variables. It has further removed unwanted factor levels in much of the data. Examine the code in the GrabData.R file to see what it is doing. Some of the variables are categorical and others are numerical. Be sure to do a variable analysis before tackling each question.  
First question - Is opinion on the death penalty (CAPPUN) independent of gun ownership (OWNGUN)?

$H_0$: Cappun is independent of owngun,they are not related  
$H_A$: They are dependent 

## Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->

CPPUN IS A categorical  variable,    
owngun is a  categorical variable,  
so i use the cat~cat data helper file    
we want CPPUN~ OWNGUN  

##Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

### Descriptive Results

#### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->



```{r}
barchartGC(~OWNGUN +CAPPUN ,data=GSSdata)
barchartGC(~OWNGUN + CAPPUN,data=GSSdata, type="percent")

```

 This graph shows that  people who own a gun are more likely to favor capital punishment
 
 
#### Numerical Descriptive Results


```{r}
table1 <- xtabs(~OWNGUN + CAPPUN,data=GSSdata)
rowPerc(table1)
colPerc(table1)
```

This shows that people who own a gun 75% are in favor then those who don't is 57% who oppose.

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->


### Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

 The p value for chi-squared test was p=0.02022.This means we reject/fail to reject the null hypothesis, and conclude that the opinion on Capital punishment (CAPPUN) is not independent of gun ownership(OWNGUN) because its less then .05
 
 
The p-value for the fisher exact test was p=0.01651 and we reject the null hypothesis as with the chi- square test. The odds ratio was 2.3 times more likely to favor CAPPUN punishment then those who don't own 
 

#  Question 2

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

Is there a difference in years of education (EDUC) between the those who did or did not vote in 2012 (VOTE12)?

$H_0$: There is not a difference in years of education between those who did or did not vote in 2012   
$H_A$:There is a difference

## Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->

 EDUC is numerical  variable   
 Vote12 is categorical variable   
 I use the file Numerical~Cat data file  
 

##Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

### Descriptive Results

#### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
plot(EDUC~VOTE12,data=GSSdata)
```

This shows that more people who voted in 2012 tend to have more education

#### Numerical Descriptive Results  

```{r}
favstats(EDUC~VOTE12,data=GSSdata)
```

This shows that people who voted12 has an average of 16 years of education and people who did not vote had 13 years of education

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

### Inferential Results

```{r}
t.test(EDUC~VOTE12,data=GSSdata)
```

The  p-value is 0.0000001 so this means we reject the null hypothesis of because the p-value is smaller then 0.05

 I conclude that there is a difference in years of education between those who did or did not vote in 2012  

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

#  Question 3

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

Determine if a persons performance on the wordsum test (WORDSUM) is independent of their level of schooling (DEGREE).  

 $H_0$: The score of a persons wordsum test is independent of their level of schooling  
 $H_A$: They are dependent  

## Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!--> 

WORDSUM is a numerical variable,  
DEGREE is a categorical variable,  
so I use the num~cat data file
 
##Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

### Descriptive Results

#### Graphical Descriptive Results  

```{r}
boxplot(WORDSUM~DEGREE,data=GSSdata)
```
 
 This shows that there is a difference between the performance of a persons test and their levels of schooling 
 
<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->

#### Numerical Descriptive Results

```{r}
favstats(WORDSUM~DEGREE,data=GSSdata)
```
 
 
  This shows that there is difference between lt highschool and the graduates because lt highschool is at a 7 and graduates at a 10
<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

### Inferential Results

```{r}
anova=aov(WORDSUM~DEGREE,data=GSSdata) 
summary(anova)
```
  This shows that we reject the null because the p-value is 0.0000002 which is smaller then .05
  
```{r}
#we specify the margins starting at the bottom and proceeding clockwise - thus 8.7 is the left side margin to hold the names of the factors
par(mar = c(7, 8.7, 2, 2))
plot(TukeyHSD(anova),las=2)
```

there is no difference between high school and less than high school. 
There is a difference with jr college and high school.
there is a difference with bachelors and high school  
there is a difference between graduate and high school  
there is no difference between high schools and graduate   
No difference between jr college and graduate   
No difference between bachelors and graduates   

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->


# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->

 
#### Question 1: is gun ownership inepedent of the opinion on capial punishment  . 
The summary results and the inferential results show that they are not independent 

#### Question 2: Is there a difference in years of education (EDUC) between the those who did or did not vote in 2012 (VOTE12)?  
In conclusion this in graphical results shows that more people who voted in 2012 tend to have more education  

#### Question 3: Determine if a persons performance on the wordsum test (WORDSUM) is independent of their level of schooling (DEGREE).
We determined in inferential results that there is a difference between the performance on the test and their level of schooling


